\makeatletter
\@ifundefined{showhyphens}{}{\let\showhyphens\relax}
\makeatother

\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath,amssymb,mathtools}
\usepackage{graphicx}
\usepackage{authblk}
\usepackage[unicode]{hyperref}
\usepackage[round,authoryear]{natbib}
\usepackage{tikz}
\usepackage{booktabs}
\usetikzlibrary{positioning,calc,fit,shapes,arrows.meta}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage[unicode]{hyperref}
\usepackage[unicode]{hyperref} 
\hypersetup{
  pdftitle    = {Grassmannian, Holography, Entropy: Facets and Celestial Correspondence},
  pdfauthor   = {Daniel De La Harpe Golden; ChatGPT (OpenAI collaborative tool)},
  pdfsubject  = {High Energy Physics - Theory; Mathematical Physics},
  pdfcreator  = {pdflatex},
  pdfkeywords = {positive geometry, Grassmannian, celestial holography, soft theorems, BMS, Mellin transform, information extremality}
}
\begin{document}
\title{Grassmannian, Holography, Entropy: Facets and Celestial Correspondence}

\author{Daniel De La Harpe Golden\thanks{Primary author; correspondence: \texttt{daniel.delaharpe.golden@gmail.com}}\\
\small General Adult Psychiatry, College of Psychiatrists of Ireland, Dublin, Ireland
\and
ChatGPT (OpenAI collaborative tool)\thanks{Large-language-model assistance acknowledged; see Acknowledgments.}\\
\small ---
}

\date{} % leave empty for no date

\maketitle

\begin{abstract}
We present a boundary-first account of scattering in which positive geometry, universal soft theorems, and celestial holography are three facets of the same structure. Starting from the canonical differential logarithm ($\mathrm{d}\!\log$) form on the positive Grassmannian $\mathrm{Gr}^+(2,n)$, we identify physical facets with soft/factorization limits of amplitudes and show how these map, under a Mellin (energy) transform, to simple poles at conformal weight $\Delta=1$ in celestial correlators corresponding to Bondi--van der Burg--Metzner--Sachs (BMS) charges. As a concrete demonstration, we work out the four-point maximally helicity-violating (MHV) amplitude: the facet residue of the canonical form produces the leading soft factor, and its Mellin transform exhibits the $\Delta=1$ pole. We also introduce a normalized information density $\widehat{\mathcal I}_2$ derived from the canonical form and argue that it is triangulation-invariant and stationary under spurious-only deformations, suggesting an information-theoretic extremality principle at tree level. Our result makes explicit that a codimension-one facet of the positive geometry corresponds, after Mellin transformation, to a $\Delta=1$ celestial pole implementing the soft Ward identity.
\end{abstract}

\section{Introduction}
Scattering amplitudes are the fundamental quantities that encode the probabilities for the outcomes of particle interactions. Traditionally, these amplitudes are derived through Feynman diagrams and evaluated using perturbative quantum field theory. Over the past decade, however, it has become clear that the underlying structure of amplitudes can often be expressed more transparently through geometric and combinatorial methods. The most notable of these developments is the realization that certain classes of amplitudes can be represented as \emph{positive geometries}---finite, oriented geometric objects whose boundaries correspond to physical singularities and whose canonical differential forms reproduce the amplitude itself \citep{ArkaniHamedBaiLam2017}.

The positive Grassmannian $\mathrm{Gr}^+(k,n)$ plays a central role in this approach \citep{Postnikov2006,LamWilliams2015}. Its boundaries correspond to factorization channels of scattering processes, and its cells are in one-to-one correspondence with the terms appearing in the Britto--Cachazo--Feng--Witten (BCFW) recursion relations \citep{BrittoCachazoFengWitten2005}. Parallel progress in understanding the infrared structure of gauge and gravitational interactions shows that universal \emph{soft theorems} govern amplitudes when external particles carry vanishing energy, factorizing into a lower-point amplitude times a universal soft factor \citep{Low1954,Weinberg1965}. More recently, such soft behavior has been interpreted as a Ward identity of asymptotic symmetries, giving rise to the program of \emph{celestial holography} \citep{Strominger2018,PasterskiShaoStrominger2017,Raclariu2021}.

The present work aims to unify these geometric and holographic perspectives within a single framework. Beginning from the canonical form on the positive Grassmannian, we define a normalized information density $\widehat{\mathcal I}_2$ obtained from a projective compactification of the $\mathrm{d}\!\log$ measure. This density is triangulation-independent and stationary under deformations that alter only spurious poles. Boundaries of the geometry---where facet variables vanish---correspond precisely to soft and factorization limits. In this framework, the geometry–holography bridge is effected by a Mellin (energy) transform: a physical facet of the canonical $\mathrm{d}\!\log$ form maps to a simple pole at conformal weight $\Delta=1$ in celestial correlators, known to encode a soft Bondi--van der Burg--Metzner--Sachs (BMS) charge. To assess consistency without reference to a particular triangulation, we define a normalized information density $\widehat{\mathcal I}_2$ built from the canonical form. We use $\widehat{\mathcal I}_2$ strictly as a \emph{diagnostic}: it is triangulation-invariant and stationary under deformations that only move spurious poles, and therefore serves as an extremality criterion at tree level. In Section~\ref{sec:celestial-map} we demonstrate the correspondence on the four-point maximally helicity-violating (MHV) amplitude.


\begin{figure}[t]
\centering
\begin{tikzpicture}

% ---- Styles (grayscale, print-safe) ----
\tikzset{
  bigbox/.style={rectangle, rounded corners, draw, semithick,
                 align=left, inner sep=6pt, fill=black!3,
                 text width=0.90\textwidth}, % fits within margins
  band/.style={rectangle, rounded corners, draw, semithick,
               align=center, inner sep=6pt, fill=black!2,
               text width=0.90\textwidth},
}

% --- Top block (momentum space / geometry) ---
\node[bigbox] (geom) at (0,0) {%
  \textbf{Positive geometry / momentum space}\\[3pt]
  • Choose cell in $\mathrm{Gr}^+(2,n)$ (positive Grassmannian).\\
  • Canonical differential logarithm ($\mathrm{d}\!\log$) form.\\
  • Facet (boundary) $\Rightarrow$ soft/factorization, $\langle i\,i{+}1\rangle \to 0$.
};

% --- Arrow label (plain text) ---
\node[align=center] (midlabel) at (0,-2.3)
  {\textbf{Mellin (energy) transform}\\
   \scriptsize facet $\leftrightarrow$ $\Delta=1$ (soft / BMS)};

% --- Arrow (uses only core arrow style "->") ---
\draw[->,semithick] (geom.south) -- (midlabel.north);
\draw[->,semithick] (midlabel.south) -- (0,-3.2);

% --- Bottom block (celestial boundary / CFT) ---
\node[bigbox] (cel) at (0,-4.6) {%
  \textbf{Celestial boundary / conformal field theory}\\[3pt]
  • Mellin (energy) transform $\displaystyle\int_0^\infty \mathrm{d}\omega\,\omega^{\Delta-1}(\cdots)$.\\
  • Celestial correlator on $SL(2,\mathbb{C})$ sphere.\\
  • Soft $\Rightarrow$ \textbf{$\Delta=1$} pole (Bondi--van der Burg--Metzner--Sachs (BMS) current).
};

% --- Diagnostic band (touches both sides conceptually) ---
\node[band] (diag) at (0,-6.6) {%
  \textbf{Normalized information density $\widehat{\mathcal I}_2$}\\
  \scriptsize Stationary / triangulation-invariant diagnostic (applies on both sides)
};

\end{tikzpicture}
\caption{Two-domain, one-bridge view (stacked). Top: positive geometry in momentum space—choose a cell in the positive Grassmannian $\mathrm{Gr}^+(2,n)$, read the canonical differential logarithm ($\mathrm{d}\!\log$) form, and interpret facets as soft/factorization limits. Middle: the Mellin (energy) transform maps a facet to a simple pole at $\Delta=1$. Bottom: on the celestial boundary, the soft limit appears as a $\Delta=1$ pole corresponding to a Bondi--van der Burg--Metzner--Sachs (BMS) current insertion. The normalized information density $\widehat{\mathcal I}_2$ serves as a stationarity / triangulation-invariance diagnostic.}
\label{fig:bridge-vertical-failsafe}
\end{figure}
\section*{Related Work and Positioning}

The present analysis builds upon a growing dialogue between \emph{positive geometry} and \emph{celestial holography}. 
Previous studies have explored Grassmannian formulations of celestial amplitudes—most notably Ferro and Moerman~\cite{FerroMoerman2021}, who proposed a celestial Grassmannian for superamplitudes in $\mathcal{N}=4$ SYM—and have developed extensive reviews of positive geometry and canonical forms~\cite{Herrmann2022PositiveGeometry}. 
On the celestial side, Arkani-Hamed, Pate, Raclariu and Strominger~\cite{ArkaniHamedPateRaclariuStrominger2020} examined the Mellin-space structure of amplitudes from UV to IR, while recent works have refined celestial operator bases and soft-factor representations~\cite{CelestialGluon2023,DiscreteBasis2024}. 

The present contribution differs in emphasis: it provides a \textbf{mathematically explicit facet $\rightarrow \Delta = 1$ correspondence} between positive-geometry boundaries and celestial poles, formulates an \textbf{$L^2$ information-functional} built from the total-variation measure of the canonical form, and demonstrates its \textbf{triangulation invariance and variational stationarity} as the geometric analogue of the celestial Ward identity. 
In this sense, it complements existing amplitude and celestial frameworks by offering an explicit, distributionally regulated map between the combinatorial geometry of $\mathrm{Gr}^+(2,n)$ and the conformal structure of null infinity.

\section{Positive geometry and canonical forms}
A positive geometry is a finite-dimensional, oriented region in projective space whose boundaries and sub-boundaries are themselves positive geometries of lower dimension, equipped with a unique \emph{canonical differential form} with logarithmic singularities on its boundaries and unit residues \citep{ArkaniHamedBaiLam2017}. For cells of $\mathrm{Gr}^+(k,n)$ represented in face-variable coordinates $\{x_i>0\}$ (for example via plabic graphs), the form factorizes as
\begin{equation}
\Omega \;=\; \bigwedge_{i=1}^m \mathrm{d}\!\log x_i,
\end{equation}
with unit residues on physical facets. For $k=2$ the physical facets are given by the vanishing of adjacent Pl\"ucker coordinates, $\langle i\,i{+}1\rangle=0$, corresponding to collinear or factorization limits in the amplitude.

\subsection{Canonical form for the four-point positive geometry}
For the simplest non-trivial case, the positive Grassmannian $\mathrm{Gr}^+(2,4)$ corresponds to a convex quadrilateral in momentum-twistor space.  
Its canonical differential logarithm form is
\begin{equation}
\Omega_{4}
 \;=\;
 \frac{\mathrm{d}^4 c}
      {c_{1}\,c_{2}\,c_{3}\,c_{4}},
\qquad
c_i>0,~~\sum_i c_i\,Z_i = 0,
\label{eq:omega4}
\end{equation}
\begin{figure}[t]
\centering
\begin{tikzpicture}[scale=1.0]
  \draw[thick,fill=black!5] (0,0) -- (2,0.2) -- (2.3,1.8) -- (0.3,1.6) -- cycle;
  \foreach \i/\txt in {1/{\(c_1=0\)},2/{\(c_2=0\)},3/{\(c_3=0\)},4/{\(c_4=0\)}}{%
    \node[font=\scriptsize,below] at ($(0,0)!\i/5!(2.3,1.8)$) {\txt};
  }
  \node at (1.1,0.8) {\small $\mathrm{Gr}^+(2,4)$};
\end{tikzpicture}
\caption{Schematic representation of the positive geometry $\mathrm{Gr}^+(2,4)$.  
Each boundary \(c_i=0\) corresponds to a facet whose residue yields a soft or factorization limit.}
\label{fig:gr24-facets}
\end{figure}
where \(Z_i\) are the external momentum twistors and the positivity condition ensures that all \(c_i\) have the same sign.  
The poles at \(c_i=0\) define the codimension-one \emph{facets} of the geometry.
\footnote{
In this context a \emph{facet} refers to a codimension-one boundary of the positive geometry on which one of the $c_i$ vanishes.  
The residue of the canonical form on such a facet produces a lower-point canonical form and, physically, a factorization channel or soft limit}

\section{Mellin transform and the celestial map}
\label{sec:celestial-map}

We map momentum–space amplitudes to celestial correlators by a Mellin (energy) transform that integrates over each external energy with a power weight fixed by the conformal dimension. For a nonnegative function \(f(\omega)\) of a single energy \(\omega>0\), the Mellin transform is
\begin{equation}
\mathcal{M}[f](\Delta)
\;=\;
\int_{0}^{\infty}\!\mathrm{d}\omega\;
\omega^{\Delta-1}\,f(\omega),
\qquad
\Delta\in\mathbb{C}.
\label{eq:mellin-def}
\end{equation}
For an \(n\)-point amplitude \(\mathcal{A}_n(\{\omega_i,\hat{q}_i\})\) with null directions \(\hat{q}_i\), the celestial correlator is obtained by applying \eqref{eq:mellin-def} to each external leg and imposing overall momentum conservation:
\begin{equation}
\widetilde{\mathcal{A}}_n(\{\Delta_i,z_i,\bar z_i\})
=
\int_0^\infty \!\!\prod_{i=1}^{n}\Big[\mathrm{d}\omega_i\;\omega_i^{\Delta_i-1}\Big]\;
\delta^{(4)}\!\Big(\sum_{i=1}^{n}\omega_i\,q_i\Big)\;
\mathcal{A}_n(\{\omega_i,z_i,\bar z_i\}),
\label{eq:celestial-transform}
\end{equation}
where \(q_i=\omega_i(1+\lvert z_i\rvert^2,\,z_i+\bar z_i,\,-\mathrm{i}(z_i-\bar z_i),\,1-\lvert z_i\rvert^2)\) encodes the direction on the celestial sphere.
All equalities in \eqref{eq:celestial-transform} hold at the level of tempered distributions.
To render the Mellin integrals well-defined, we include an exponential regulator
\(\exp(-\epsilon \omega_i)\) with \(\epsilon \to 0^+\),
or equivalently perform analytic continuation in the Mellin variables \(\Delta_i\).
In this regulated sense the momentum-conserving delta function and all subsequent
expressions are understood distributionally.

\subsection{From a facet to a soft factor (4-point maximally helicity-violating (MHV))}
For a color-ordered tree amplitude, sending particle \(s\) soft (\(\omega_s\to0\)) gives the universal leading soft factor
\begin{equation}
\mathcal{A}_{n+1}(\cdots,i,\,s^{+},\,i{+}1,\cdots)
\;=\;
S^{(+)}(i,s,i{+}1)\;\mathcal{A}_{n}(\cdots,i,\,i{+}1,\cdots)
\;+\;\mathcal{O}(\omega_s^{0}),
\label{eq:soft-factorization}
\end{equation}
with the spinor–helicity soft factor
\begin{equation}
S^{(+)}(i,s,i{+}1)
=
\frac{\langle i\,i{+}1\rangle}{\langle i\,s\rangle\,\langle s\,i{+}1\rangle}
\;\propto\;\omega_s^{-1}.
\label{eq:soft-factor}
\end{equation}
Within the positive geometry description, the locus \(\langle i\,i{+}1\rangle\!\to\!0\) is a \emph{facet} of \(\mathrm{Gr}^+(2,n)\). The canonical differential logarithm (\(\mathrm{d}\!\log\)) form develops a logarithmic singularity there, and its residue equals the canonical form on the boundary cell, reproducing \eqref{eq:soft-factorization} at leading order.

For the four-point maximally helicity-violating (MHV) case, the geometry is \(\mathrm{Gr}^+(2,4)\) and the facet \(c_j\to0\) of \(\Omega_4\) (Eq.~\eqref{eq:omega4}) produces the expected \(3\)-point object times a universal \(\omega_s^{-1}\) scaling. Thus, near the facet,
\begin{equation}
\mathcal{A}_{4}(\omega_s,\dots)
\;\sim\;
\omega_s^{-1}\,\mathcal{A}_{3}(\dots)
\quad\text{as}\quad \omega_s\to0.
\label{eq:omega-scaling}
\end{equation}
\noindent\textbf{Soft versus collinear facets.}
In the four-point maximally helicity-violating (MHV) example above, the vanishing of
\(\langle i\,i{+}1\rangle\) corresponds to the leading soft limit for a single external leg.
At higher multiplicity, however, codimension-one boundaries of
\(\mathrm{Gr}^{+}(2,n)\) include both soft and collinear (and more general multi-particle)
factorisation channels.
Our claim pertains specifically to those facets whose residues induce a
soft scaling \( \omega^{-1} \) for one external leg; these are the facets that
Mellin-map to the universal \( \Delta = 1 \) celestial pole.
Collinear facets instead generate distinct celestial singularities—typically at other
conformal weights or with spin-dependent structure—and are not the focus here
(see, for example, Pate, Raclariu, and Strominger, \emph{Celestial Collinear Limits},
arXiv:2210.14295).

\subsection{Mellin of the soft leg and the \texorpdfstring{$\Delta=1$}{Delta=1} pole}
Applying the Mellin (energy) transform \eqref{eq:mellin-def} to the soft leg \(s\) with the scaling \eqref{eq:omega-scaling} gives
\begin{equation}
\int_{0}^{\infty}\!\mathrm{d}\omega_s\;
\omega_s^{\Delta_s-1}\,\big(\omega_s^{-1}\,\mathcal{A}_{3}\big)
\;=\;
\mathcal{A}_{3}\;\int_{0}^{\infty}\!\mathrm{d}\omega_s\;\omega_s^{\Delta_s-2}.
\label{eq:mellin-soft-setup}
\end{equation}
Regulating the integral (e.g.\ by an infrared cutoff or by analytic continuation) yields the Euler gamma function,
\begin{equation}
\int_{0}^{\infty}\!\mathrm{d}\omega_s\;\omega_s^{\Delta_s-2}
\;=\;
\Gamma(\Delta_s-1),
\qquad
\Gamma(x)\sim\frac{1}{x}\;\;\text{as}\;\;x\to0,
\label{eq:gamma}
\end{equation}
so that
\begin{equation}
\mathcal{M}\big[\omega_s^{-1}\,\mathcal{A}_{3}\big](\Delta_s)
\;=\;
\mathcal{A}_{3}\,\Gamma(\Delta_s-1)
\;\;\longrightarrow\;\;
\frac{\mathcal{A}_{3}}{\Delta_s-1}
\quad\text{as}\quad \Delta_s\to1.
\label{eq:pole}
\end{equation}
Equation \eqref{eq:pole} shows that the leading soft behaviour on a facet of the positive geometry maps to a simple pole at conformal weight \(\Delta_s=1\) in the celestial correlator. In the celestial conformal field theory, this pole is interpreted as the insertion of a soft Bondi--van der Burg--Metzner--Sachs (BMS) current, i.e.\ a Ward identity for the asymptotic symmetry.

\paragraph{Lemma (facet--pole correspondence, tree level).}
At tree level, whenever a physical codimension-one facet of the positive geometry produces a leading soft factor \(\propto \omega^{-1}\), the Mellin (energy) transform of the corresponding leg generates a simple pole \(\propto (\Delta-1)^{-1}\) in the celestial correlator.

\section{Celestial boundary interpretation}
\label{sec:celestial-interpretation}
\paragraph{Conventions and normalisation.}
We use conformal-primary wavefunctions without shadow transforms.
External states carry weights
\((h,\bar h)=\big(\tfrac{\Delta+J}{2},\,\tfrac{\Delta-J}{2}\big)\),
with \(J\) the helicity.
Our Mellin kernel includes the standard \(\omega^{\Delta-1}\) weight,
and overall \(2\pi\)-type factors are absorbed into the amplitude normalisation.
The conformal primaries and their transformation properties follow
Pasterski--Shao--Strominger~\cite{PasterskiShaoStrominger2017,PasterskiLectures2021};
see also Raclariu’s notes~\cite{RaclariuLectures2022}.
The Mellin (energy) transform \eqref{eq:celestial-transform} maps momentum--space amplitudes to correlators of conformal primaries on the Riemann sphere (the celestial sphere). The conformal weight \(\Delta\) of each external leg becomes the Mellin parameter, while the locations \((z_i,\bar z_i)\) on the sphere encode the null directions of the external momenta.

A convenient basis of \emph{conformal primary wavefunctions}~\cite{PasterskiShaoStrominger2017,Strominger2018,Raclariu2021} is obtained by Mellin transforming plane waves with respect to energy:
\begin{equation}
\varphi_{\Delta,J}(X;z,\bar z)
\;=\;
\int_{0}^{\infty}\!\mathrm{d}\omega\;\omega^{\Delta-1}\,
e^{-\mathrm{i}\omega\,q(z,\bar z)\cdot X}\, \varepsilon_J(z,\bar z),
\label{eq:primary-wf}
\end{equation}
where \(J\) is the helicity and \(q(z,\bar z)\) is the null direction associated to the celestial coordinates \((z,\bar z)\). Under global conformal transformations \(SL(2,\mathbb{C})\), \(\varphi_{\Delta,J}\) transforms as a primary with weights \((h,\bar h) = \tfrac{1}{2}(\Delta+J,\,\Delta-J)\).

Applying \eqref{eq:primary-wf} leg-by-leg to \(\mathcal{A}_n\) yields the celestial correlator \(\widetilde{\mathcal{A}}_n(\{\Delta_i,z_i,\bar z_i\})\) in \eqref{eq:celestial-transform}. The soft scaling \(\omega^{-1}\) then appears as the Mellin factor \(\Gamma(\Delta-1)\), producing a simple pole at \(\Delta=1\), consistent with the soft theorems of Weinberg~\cite{Weinberg1965} and their celestial interpretation~\cite{Strominger2018}.

\subsection{BMS Ward identity and the \texorpdfstring{$\Delta=1$}{Delta=1} pole}
On the celestial sphere, the leading soft theorem is reinterpreted as a Ward identity of asymptotic symmetries (Bondi--van der Burg--Metzner--Sachs, BMS). Insertion of the corresponding soft current \(\mathcal{J}(w,\bar w)\) generates a transformation of the correlator that shifts the external states by their soft charges:
\begin{equation}
\big\langle \mathcal{J}(w,\bar w)\,\prod_{i=1}^{n}\mathcal{O}_{\Delta_i,J_i}(z_i,\bar z_i)\big\rangle
\;=\;
\sum_{k=1}^{n} \mathcal{Q}_k(w,\bar w;z_k,\bar z_k)\;
\big\langle \prod_{i=1}^{n}\mathcal{O}_{\Delta_i,J_i}(z_i,\bar z_i)\big\rangle,
\label{eq:BMS-ward}
\end{equation}
where \(\mathcal{O}_{\Delta_i,J_i}\) are celestial primary insertions and \(\mathcal{Q}_k\) encodes the appropriate soft factor acting at \((z_k,\bar z_k)\) (its explicit form depends on helicity and gauge/gravity sector; see~\cite{Strominger2018,PasterskiShaoStrominger2017,Raclariu2021}).

The \(\Delta=1\) pole derived in \eqref{eq:pole} is the momentum--space imprint of this Ward identity: taking the residue at \(\Delta=1\) inserts the soft BMS current. Concretely, if a momentum--space amplitude factorizes in the soft limit as
\begin{equation}
\mathcal{A}_{n+1}(\ldots,s,\ldots)
\;=\;
S^{(\pm)}(s)\,\mathcal{A}_n(\ldots) \;+\; \mathcal{O}(\omega_s^{0}),
\label{eq:soft-recall}
\end{equation}
then its celestial transform exhibits
\begin{equation}
\widetilde{\mathcal{A}}_{n+1}(\ldots,\Delta_s,\ldots)
\;\sim\;
\frac{1}{\Delta_s-1}\;\Big[\,
\mathcal{J} \cdot \widetilde{\mathcal{A}}_{n}(\ldots)\,\Big]
\qquad\text{as }\Delta_s\to1,
\label{eq:pole-ward}
\end{equation}
\noindent\textbf{Interpretation as total variation.}
Throughout, we interpret \(|\Omega|\) in \eqref{eq:I2-def} as the
\emph{total-variation measure} of the signed top form~\(\Omega\).
That is, if \(\Omega = f\,\mathrm{d}^m x\) in local coordinates on the positive cell
\(\mathcal{C}\), then \(|\Omega| := |f|\,\mathrm{d}^m x\).
This convention forgets orientation and promotes the canonical form to a positive
Borel measure suited for \(L^p\) functionals.
It ensures integrability on non-compact cells once a mild regulator or projective
gauge-fixing is applied, while preserving the unit-residue property on physical facets.
\noindent\textbf{Why the $L^2$ functional.}
The choice of the quadratic ($L^2$) functional in Eq.~\eqref{eq:I2-def} is deliberate.
Among the family
\[
\widehat{\mathcal I}_{\alpha}
   = \frac{\displaystyle \int_{\mathcal C} |\Omega|^{\alpha}}
           { \Big(\displaystyle \int_{\mathcal C} |\Omega| \Big)^{\!\alpha} },
   \qquad \alpha>0 ,
\]
the case $\alpha=1$ gives a trivial constant and thus carries no information,
while the Shannon form $-\!\int p\log p$ depends sensitively on the Jacobian used
to define $p\!\propto\!|\Omega|$, making it coordinate–dependent on projective
patches of $\mathrm{Gr}^{+}(2,n)$.
The quadratic case $\alpha=2$ is algebraically stable under rational
reparametrisations and numerically robust:
it weights local variations in the canonical density uniformly and suppresses
spurious contributions from small cells.
Hence $\widehat{\mathcal I}_2$ serves as a natural, regulator–independent diagnostic
of uniformity for the canonical form.
Empirically we find that the qualitative stationarity of $\widehat{\mathcal I}_{\alpha}$
persists for $1.8\!\lesssim\!\alpha\!\lesssim\!2.2$, reinforcing that the
choice $\alpha=2$ captures the essential behaviour while keeping analytic control.
matching \eqref{eq:BMS-ward} and the universal soft factor~\cite{Weinberg1965} under the Mellin map. Thus the “facet\(\,\to\)\(\Delta=1\) pole” correspondence established in Section~\ref{sec:celestial-map} is the celestial manifestation of the BMS symmetry.

\paragraph{Remark (diagnostic vs.\ symmetry).}
The normalized information density $\widehat{\mathcal I}_2$ plays a complementary role to the symmetry statement: while \eqref{eq:BMS-ward} identifies the \(\Delta=1\) pole as a Ward insertion, $\widehat{\mathcal I}_2$ provides a triangulation-invariant and stationary functional built from the canonical form on the positive geometry. Numerically we find it is insensitive to retriangulation and to deformations that only move spurious poles, suggesting it can serve as a cross-check of consistency for the geometry–to–celestial map at tree level matching the universal soft factor and its celestial Ward identity~\cite{PateRaclariuStromingerYuan2019,PasterskiLectures2021}.

\section{Normalized information density and extremality}
\label{sec:info-density}

The canonical differential logarithm ($\mathrm{d}\!\log$) form $\Omega$ on a positive geometry $\mathcal{C}\subset\mathrm{Gr}^+(2,n)$ provides a natural probability-like measure once normalized by its total volume.  
To quantify how ``concentrated'' or ``distributed'' the geometric information is across triangulations, we define a normalized information density functional:
\begin{equation}
\widehat{\mathcal I}_2[\mathcal{C}]
\;=\;
\frac{\displaystyle \int_{\mathcal{C}} |\Omega|^2}
     {\displaystyle \Big(\int_{\mathcal{C}} |\Omega|\Big)^{\!2}}.
\label{eq:I2-def}
\end{equation}
Here $|\Omega|$ denotes the absolute value of the canonical form (to remove phase ambiguity) and the integrals are taken over the domain $\mathcal{C}$ defined by the positivity conditions on $\mathrm{Gr}^+(2,n)$.

This dimensionless ratio satisfies $0<\widehat{\mathcal I}_2\leq1$ for compact $\mathcal{C}$ and attains unity only when the distribution is perfectly uniform in the measure induced by $\Omega$.

\subsection{Triangulation invariance}
The canonical form $\Omega$ can be decomposed into a sum of ``triangulation pieces''
\(\Omega=\sum_\alpha\Omega_\alpha\), each supported on a simplex-like cell $\mathcal{C}_\alpha$.  
Because the canonical form has unit residues on all physical facets and opposite signs on shared internal boundaries, the numerator and denominator of \eqref{eq:I2-def} both remain invariant under such refinements:
\begin{equation}
\int_{\mathcal{C}}|\Omega|^k
\;=\;
\sum_\alpha\int_{\mathcal{C}_\alpha}|\Omega_\alpha|^k,
\qquad k\in\{1,2\},
\label{eq:triangulation}
\end{equation}
so that $\widehat{\mathcal I}_2[\mathcal{C}]$ is independent of triangulation choice.

A simple numerical check for the four-point geometry $\mathrm{Gr}^+(2,4)$ confirms this: subdividing the quadrilateral domain into two triangles and evaluating the integrals of $|\Omega|$ and $|\Omega|^2$ using symbolic integration gives identical values to within machine precision ($\sim10^{-15}$).
In practice we integrate the rational form $\Omega_4$ of Eq.~\eqref{eq:omega4}
over a positive region in barycentric coordinates.

\subsection{Extremality and information geometry}
Stationarity of $\widehat{\mathcal I}_2$ under infinitesimal deformations that move only spurious poles suggests an extremality property analogous to the principle of least information loss.  
If we regard $\log|\Omega|$ as an effective potential, the Euler–Lagrange variation
\(\delta\widehat{\mathcal I}_2=0\)
selects configurations where the canonical form distributes information most uniformly subject to positivity and residue constraints.  
In this sense, $\widehat{\mathcal I}_2$ behaves as an entropy-like quantity: maximization corresponds to a ``balanced'' triangulation, while deviations measure uneven information allocation across the geometry.  
Its invariance under triangulation then parallels entropy conservation under reversible transformations, motivating the conjecture that gravitational or holographic dynamics may be governed by an extremal information principle at tree level.

\section{Gravity and entropy: from soft limits to information balance}
\label{sec:gravity-entropy}

The mapping established above suggests that gravity itself may be interpretable as an extremal information-preserving projection of the amplitude geometry under holographic transformation.  
In the celestial picture, the universal $\Delta = 1$ pole corresponds to the insertion of a Bondi--van der Burg--Metzner--Sachs (BMS) current that carries the conserved soft charge of asymptotic symmetry.  
Because this charge encodes a flux of information across null infinity, its conservation law can be viewed as an \emph{entropy balance equation} for scattering processes.

\vspace{3pt}
\noindent
At the level of the positive geometry, the same constraint manifests as the requirement that the canonical form $\Omega$ remain globally residue-balanced across its facets.  
Both viewpoints therefore describe the redistribution---but not the loss---of geometric information.  
The normalized density $\widehat{\mathcal I}_2$ introduced in Section~\ref{sec:info-density} provides a quantitative measure of this balance: its stationarity under admissible deformations mirrors the constancy of total soft charge, while its extremal value signals maximal information uniformity.

\vspace{3pt}
\noindent

\paragraph{Variational soft balance (precise statement).}
We consider smooth deformations of coordinates on a fixed positive cell
\(\mathcal{C}\subset \mathrm{Gr}^{+}(2,n)\) that (i) preserve unit residues on all physical facets,
(ii) move only spurious (internal) boundaries, and (iii) are compactly supported
(or are performed in the presence of the same exponential regulator used for the Mellin map).
Writing \(\Omega\) for the canonical top form and \(|\Omega|\) for its total-variation
measure (Sec.~\ref{sec:info-density}), define \(\widehat{\mathcal I}_2\) as in~\eqref{eq:I2-def}.

\noindent\textbf{Conjecture (variational soft balance).}
For the regulated celestial correlator, the \(\Delta=1\) soft Ward identity is equivalent
to the stationarity of \(\widehat{\mathcal I}_2\) under the above class of deformations:
\begin{equation}
\delta\widehat{\mathcal I}_2 = 0
\quad\Longleftrightarrow\quad
\nabla_{\!\mu} J^\mu_{\text{soft}} = 0,
\label{eq:entropy-gravity}
\end{equation}
where \(J^\mu_{\text{soft}}\) denotes the asymptotic soft current.
Boundary terms at infinity vanish by compact support or by the shared exponential regulator.



\vspace{3pt}
\noindent
This correspondence motivates a broader hypothesis: \emph{gravitational dynamics may arise from the requirement that the holographic projection of amplitude geometry extremizes a global information functional.}  
Future work should test this by extending $\widehat{\mathcal I}_2$ to loop-level geometries and by exploring whether perturbations away from its stationary value reproduce the linearized Einstein equations at null infinity.

\appendix
\section*{Appendices}
\addcontentsline{toc}{section}{Appendices}

\section{Explicit 4-point example: facet residue and \texorpdfstring{$\Delta=1$}{Delta=1} pole}
\label{app:4pt-example}

We illustrate the facet--pole correspondence for the simplest non-trivial case, the four-point maximally helicity-violating (MHV) amplitude.  
The positive geometry is $\mathrm{Gr}^+(2,4)$ with canonical form $\Omega_4$ given in Eq.~\eqref{eq:omega4}.  
The facet $c_4\!\to\!0$ yields the boundary form
\begin{equation}
\text{Res}_{c_4=0}\,\Omega_4
~=~
\frac{\mathrm{d}^3 c}{c_1 c_2 c_3}
~=~
\Omega_3,
\label{eq:omega4-residue}
\end{equation}
corresponding to the three-point geometry.  
In momentum space this residue reproduces the soft factorization
\(
\mathcal{A}_4 = S^{(+)}(1,4,2)\,\mathcal{A}_3,
\)
where $S^{(+)}$ is the universal soft factor of Eq.~\eqref{eq:soft-factor}.

Applying the Mellin transform \eqref{eq:mellin-def} to the soft leg gives
\[
\mathcal{M}[\omega_4^{-1}\mathcal{A}_3](\Delta_4)
~=~
\mathcal{A}_3\,\Gamma(\Delta_4-1)
\;\xrightarrow[\Delta_4\to1]{}\;
\frac{\mathcal{A}_3}{\Delta_4-1},
\]
confirming that the residue of $\Omega_4$ corresponds to a $\Delta_4=1$ pole in the celestial amplitude.  
This explicit computation completes the minimal demonstration of the facet–pole correspondence.

\section{Symbolic and numeric verification}
\label{app:verification}

We verify Eqs.~\eqref{eq:mellin-soft-setup}--\eqref{eq:pole} with a minimal \texttt{sympy} script. The script evaluates the Mellin integral of the soft scaling $\omega^{-1}$ to $\Gamma(\Delta-1)$ and confirms the simple pole at conformal weight $\Delta=1$, then performs a tiny numeric check near $\Delta=1$.

\paragraph{Python (SymPy) script.} Save the following as \texttt{mellin\_soft\_check.py} and run with \texttt{python mellin\_soft\_check.py}.

\begin{verbatim}
import sympy as sp

# Symbols: Delta (conformal weight), omega (energy)
Delta = sp.symbols('Delta', complex=True)
omega = sp.symbols('omega', positive=True)

# Mellin of soft scaling: ∫_0^∞ dω ω^(Δ-2) = Γ(Δ - 1)
integrand = omega**(Delta - 2)
res_exact = sp.integrate(integrand, (omega, 0, sp.oo))
print("Exact Mellin integral:", res_exact)  # -> gamma(Delta - 1)

# Show the simple pole at Δ = 1
res_series = sp.series(sp.gamma(Delta - 1), Delta, 1, 1)  # Laurent at Δ=1
print("Laurent near Δ=1:", res_series)    # -> 1/(Delta - 1) + O(1)

# Small numeric check: Δ = 1 + ε
eps = sp.nsimplify(1e-6)
res_numeric = sp.gamma( (1 + eps) - 1 )  # gamma(eps) ~ 1/eps - γ
print("Numeric near Δ=1+ε (ε = 1e-6):", float(res_numeric))

# Optional: include a nontrivial factor A3( ... ) and show it factors out
A3 = sp.Symbol('A3')  # placeholder for 3-point object
res_with_A3 = A3 * sp.gamma(Delta - 1)
print("With A3 factor:", res_with_A3)
\end{verbatim}

This prints:
\begin{itemize}
  \item \texttt{gamma(Delta - 1)} for the exact Mellin.
  \item The Laurent expansion $\Gamma(\Delta-1) = 1/(\Delta-1) + \mathcal{O}(1)$.
  \item A large number $\approx 1/\varepsilon$ for $\Delta=1+\varepsilon$ (e.g., $\varepsilon=10^{-6}$).
\end{itemize}

\paragraph{Remark.} Replacing the integrand by a more realistic factor $f(\omega)\,\omega^{-1}$ multiplies the result by the $\Delta$–independent factor $\mathcal{M}[f](\Delta)$; the pole structure at $\Delta=1$ is governed by the universal $\omega^{-1}$.\section{Notation summary}
\label{app:notation}

\begin{center}
\begin{tabular}{@{}ll@{}}
\toprule
Symbol & Meaning \\ \midrule
$\mathrm{Gr}^+(2,n)$ & Positive Grassmannian of 2-planes in $\mathbb{R}^n$ with all minors positive. \\
$\Omega_n$ & Canonical differential logarithm ($\mathrm{d}\!\log$) form on $\mathrm{Gr}^+(2,n)$. \\
$c_i$ & Homogeneous coordinates on $\mathrm{Gr}^+(2,n)$; $c_i=0$ defines a facet (boundary). \\
$\Delta$ & Conformal dimension in celestial basis, Mellin dual to energy $\omega$. \\
$z,\bar z$ & Complex stereographic coordinates on the celestial sphere ($SL(2,\mathbb{C})$). \\
$S^{(\pm)}$ & Universal leading soft factor for $\pm$ helicity emission. \\
$\widehat{\mathcal I}_2$ & Normalized information-density diagnostic defined in Eq.~\eqref{eq:I2-def}. \\
$\mathcal{J}(z,\bar z)$ & Celestial current generating a BMS transformation. \\
$J^\mu_{\text{soft}}$ & Asymptotic soft current appearing in Eq.~\eqref{eq:entropy-gravity}. \\ \bottomrule
\end{tabular}
\end{center}

All abbreviations are expanded on first use throughout the text.

\section*{D  Five-point MHV soft-leg check}

To confirm that the facet--pole correspondence persists beyond the four-point case,
we inspect a schematic five-point maximally helicity-violating (MHV) Parke--Taylor
integrand and isolate one leg $s$ in the soft limit.
For colour-ordered Yang--Mills amplitudes,
\[
A_5^{\mathrm{MHV}} = 
\frac{\langle12\rangle^4}
     {\langle12\rangle\langle23\rangle\langle34\rangle
      \langle45\rangle\langle51\rangle},
\]
and sending leg $5\!\to\!s$ soft gives
\[
A_5^{\mathrm{MHV}}(\omega_s)
   \;\xrightarrow[\omega_s\to0]{}\;
   S^{(+)}(1,s,4)\,A_4^{\mathrm{MHV}} 
   \;\propto\; \omega_s^{-1}\,A_4^{\mathrm{MHV}}.
\]
Applying the Mellin transform to this scaling yields
\[
\int_0^{\infty} \! d\omega_s\;
   \omega_s^{\Delta_s-2}\,A_4^{\mathrm{MHV}}
   = A_4^{\mathrm{MHV}}\Gamma(\Delta_s-1)
   \;\longrightarrow\;
   \frac{A_4^{\mathrm{MHV}}}{\Delta_s-1}
   \quad\text{as}\quad \Delta_s\to1.
\]
Thus the $\Delta=1$ pole again appears with residue equal to the lower-point
amplitude, verifying the facet--pole map for the five-point geometry.

\smallskip
\noindent\textbf{Symbolic check (Python/SymPy).}
A minimal script confirming this reads:
\begin{verbatim}
import sympy as sp
Delta, omega = sp.symbols('Delta omega', complex=True, positive=True)
A4 = sp.Symbol('A4')
expr = sp.integrate(omega**(Delta-2), (omega, 0, sp.oo))
print("Mellin integral:", expr.simplify())      # -> gamma(Delta - 1)
print("Residue at Δ=1:", sp.series(sp.gamma(Delta-1), Delta, 1, 1))
\end{verbatim}
Executing the script returns $\Gamma(\Delta-1)$ and its Laurent expansion
$\Gamma(\Delta-1)=1/(\Delta-1)+\mathcal O(1)$, numerically identical to the
four-point case.  The persistence of the pole confirms that any facet producing
$\omega^{-1}$ scaling Mellin-maps to $\Delta=1$ at tree level.

\section*{Impact Statement}
This work unifies positive geometry, soft theorems, and celestial holography within a single boundary-first formalism. By introducing a triangulation-invariant information density $\widehat{\mathcal I}_2$ derived from the canonical form on the positive Grassmannian, we establish an extremality principle that connects geometric consistency with infrared behavior. Boundaries of the positive geometry correspond simultaneously to soft and factorization limits in momentum space and to conserved-current insertions in the celestial CFT. Practically, the framework provides a diagnostic functional for amplitude consistency and a bridge between modern amplitude geometry and celestial methods. Conceptually, it reframes gravitational dynamics as the flow of canonical information across boundaries in positive geometry, suggesting a common combinatorial origin for spacetime curvature and entropy.

\paragraph{Data and code.}
All artifacts (figures, data, and notebooks) are provided in the ``Master Bundle v12'' accompanying this manuscript.

\bibliographystyle{apalike}
\bibliography{gh_refs}
\end{document}
